### 2020-12-16 experiments/eiv_max_likelihood.Rmd
First working commit. Univariate y(t) = A*(x(t+tau)) + C

---
title: "eiv_max_likelihood"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Define the problem.

```{r definition}
library("mvtnorm")
library("tidyverse")

a = 2
constant = 2
x = runif(100,0,10)
xi = x + rnorm(100,0,1)
plot(x, xi)
# Define a cubic function
y = a*x^2 + rnorm(100,0,10) + constant
plot(xi, y)

```

```{r}
# Plot the results of the regression without considering the EIV
lm(y~I(xi^2))
```
```{r}
# Define the solver functions
all_likelihoods = function(xi, y, x_theta, y_theta, alpha=alpha, c=c){

  shifts = NULL
  sum_l = 0
  for (i in 1:length(xi)){
    #print(i)
    point_l = point_likelihood(xi[i], y[i], x_theta, y_theta, alpha=alpha, c=c)
    #print(point_l)
    sum_l = sum_l + point_l$likelihood
    shifts = append(shifts, point_l$shift_z)
  }
  # Return the total shift likelihood across all points
  return(sum_l)
}

all_likelihoods_optim = function(params, xi, y){

  print(params)
  x_theta = params[1]
  y_theta=params[2]
  alpha=params[3]
  c=params[4]

  shifts = NULL
  sum_l = 0
  for (i in 1:length(xi)){
    #print(i)
    point_l = point_likelihood(xi[i], y[i], x_theta, y_theta, alpha=alpha, c=c)
    #print(point_l)
    sum_l = sum_l + point_l$likelihood
    shifts = append(shifts, point_l$shift_z)
  }

  print(sum_l)
  # Return the total shift likelihood across all points
  return(sum_l)
}

point_likelihood = function(xi0, yi, x_theta, y_theta, alpha=2, c=0){

  z_optim = optim(0, 
        fn=single_likelihood, 
        x_theta = x_theta,
        y_theta=y_theta, 
        yi=yi, 
        xi = xi0,
        alpha=alpha,
        c=c, 
        method="Brent",
        lower=-6*x_theta,
        upper=6*x_theta,
        type="mv")

  shift_z = z_optim$par
  max_l = z_optim$value

  return(list("shift_z"=(shift_z), "likelihood"=max_l))
}

single_likelihood = function(params, x_theta, y_theta, xi0, yi, alpha, c, type="individual"){

    z = params[1]
    xl = dnorm(z, 0, sd=x_theta, log=TRUE)
    y_p = alpha*(z + xi0)^2 + c
    res = yi - y_p
    yl = dnorm(res, 0, y_theta, log=TRUE)
    #print(paste0("XL: ",xl, " , YL: ", yl, ", Shift: ",z, ", Res: ",res))

    if (type == "individual"){
      return (-(yl + xl))
    } else {
      # Calculate the likelihood from a Multivariate Gaussian
      # Assume independence of the parameters
      covariance_mat = matrix(c(x_theta, 0, 0, y_theta), nrow=2, ncol=2)
      means = c(0,0)
      m_l = dmvnorm(c(z, res), means, covariance_mat, log = TRUE)
      return(-m_l)
    }
}
```

# Monte Carlo Simulation with Plots
```{r monte_carlo}
oy = c(1:100)/10
ox = c(1:100)/10
a = 1 + 1:20/10
c = c(1:100)/10

monte_carlo = as.data.frame(expand.grid(a, c, oy, ox))
monte_carlo$likelihood = NA
colnames(monte_carlo) = c("alpha","c","oy","ox","likelihood")
#monte_carlo$r = monte_carlo$ox/monte_carlo$oy

# Take a subset

monte_carlo = monte_carlo[sample(1:nrow(monte_carlo), 10000),]

for (i in 1:nrow(monte_carlo)){
  print(i)
  # Assume independent 
  l = all_likelihoods(xi, y, x_theta=monte_carlo[i,'ox'], y_theta=monte_carlo[i,'oy'], alpha=monte_carlo[i,'alpha'], c=monte_carlo[i,'c'])
  monte_carlo[i,'likelihood'] = l
}
```
```{r}
# Optimised version

model_fit = optim(c(1,1,1,1), fn=all_likelihoods_optim, xi=xi, y=y, method="L-BFGS")

```

```{r}
model_fit$par
```

```{r outcome}
monte_carlo = monte_carlo[order(monte_carlo$likelihood),]

plot(monte_carlo$ox, monte_carlo$likelihood, ylim=c(0,50))
```

```{r}
head(monte_carlo, 50)
plot(monte_carlo$ox, monte_carlo$likelihood, ylim=c(120, 1000))
```

```{r}
xt= (1:100)/10
y_p = 2.1*(xt)^2 - 15
plot(xt, y_p, ylim=c(-20, 250))
points(xi, y, col="red")
```

```{r}

```